---
title: "Senior Worker 450 Code"
author: "Zhen Liu, Yutong Liu, Jingyi Huang"
date: "2018/2/1"
output: html_document
---

```{r data input}
s<-getwd()
substr(s, 1, nchar(s)-4)
path<-paste(substr(s, 1, nchar(s)-4),"Data/Data-NFEHRS_revised.sav",sep = "")
library(haven)
Rawdata<- read_sav(path)  #1247
```

## Data Pre-processing
### 1. Check Response
```{r response skil_num}
summary(Rawdata$num_use)
nrow(Rawdata[Rawdata$num_use==7,])  
```
The max of num_use is 5 so I guessed 7 may be a wrong input. Also, there are other four missing value in this observation.Therefore, we'll use the dataset that exclude this obervation. observation(num_use=7):
```{r responses qqplot }
newdata<- Rawdata[Rawdata$num_use!=7,]  #1246
par(mfrow=c(2,2))
#1)skill_num
summary(newdata$num_use)
qqnorm(newdata$num_use)
qqline(newdata$num_use)

#2) skill_lit
summary(newdata$liter_use)
qqnorm(newdata$liter_use)
qqline(newdata$liter_use)

#3) test_num
summary(newdata$pvnumM)
qqnorm(newdata$pvnumM)
qqline(newdata$pvnumM)

#4) test_lit
summary(newdata$pvlitM)
qqnorm(newdata$pvlitM)
qqline(newdata$pvlitM)

```

The top left and top right Q-Q plot are not on a straight line. We might need some further transformation on the responses(num_use, liter_use)
 
The bottom left and bottom right Q-Q plot seem on a straight line so it is fair to say that the two responses(pvnumM, pvlitM) of our sample are normal distributed. The assumption of linear regression is statisfied.


### 2.Show summary statistics and check Missing Value
```{r results= 'hide'}
sum.data<-summary(newdata)[7,]
sum.data[!is.na(sum.data)]
```
There are 501 missing value in the 'Mgr' column, which is more than 40% of the total number of the observations. I would suggest that progressing our regression analysis without this variable first. 

There are 39 missing value in the 'pub_priv'. 
```{r pub_priv}
table(newdata$pub_priv)
```
From the table above, 1062 of the observations are from private sector while 145 of them are from Public sector. 

### 3.Check Variable
```{r}
##1.ED_Level
unique(newdata$ED_Level)
nrow(newdata[newdata$ED_Level == 8,]) #only one observation, changes it to 4
newdata<- within(newdata, ED_Level[ED_Level==8]<-4 )

##2.Categorical variable
newdata$pub_priv <- as.factor(newdata$pub_priv)
newdata$GENDER_R <- as.factor(newdata$GENDER_R)
newdata$ED_Level <- as.factor(newdata$ED_Level)
newdata$Full_part <- as.factor(newdata$Full_part)
newdata$NFE12 <- as.factor(newdata$NFE12)
newdata$FNFE12JR <- as.factor(newdata$FNFE12JR)
newdata$FNFAET12NJR <- as.factor(newdata$FNFAET12NJR)
newdata$FNFAET12JR <- as.factor(newdata$FNFAET12JR)

##3. "FNFAET12" "FNFAET12JR" "FNFAET12NJR" t-test:
#1)pvlitM
t.test(pvlitM~FNFAET12, data = newdata)$p.value #p-value < 2.2e-16
t.test(pvlitM~FNFAET12JR, data = newdata)$p.value #p-value = 9.813e-13
t.test(pvlitM~FNFAET12NJR, data = newdata)$p.value #p-value = 0.0001315

#2)pvnumM
t.test(pvnumM~FNFAET12, data = newdata)$p.value #p-value < 2.2e-16
t.test(pvnumM~FNFAET12JR, data = newdata)$p.value #p-value < 2.2e-16
t.test(pvnumM~FNFAET12NJR, data = newdata)$p.value #p-value = 6.66e-06
##they are all statistically significant and FNFAET12 is the union of other two variables.

##4.FNFE12JR and FNFAET12JR
table(newdata$FNFE12JR == newdata$FNFAET12JR) #They are exactly the same: use one of them

##Two dataset:
#1) proficiency test score(numeracy) + possbile factor
data_num<- newdata[,c(1,2,5,7,9,10,13,14,17,19,20,23,24,25)]
data_num<- na.omit(data_num) #1207
#2) proficiency test score(literacy) + possbile factor
data_lit<- newdata[,c(1,2,5,7,9,10,13,14,17,18,20,23,24,25)]
data_lit<- na.omit(data_lit) #1207


```

### 4. Correlations
```{r}
newdata.numeric<-newdata[,c("AGE_R","Years_wk","work_flexM","work_lrnM","act_lrn","NFEHRS","pvlitM","pvnumM")]
m<-cor(newdata.numeric)
if (!require('corrplot')) install.packages('corrplot')
library('ggplot2')
corrplot(m, method = "circle") #plot matrix

categorical<-c("GENDER_R","Full_part","ED_Level","pub_priv","NFE12","FNFAET12JR","FNFAET12NJR")
catcorrm <- function(categorical, newdata) sapply(categorical, function(y) sapply(categorical, function(x) assocstats(table(unlist(newdata[,x]), unlist(newdata[,y])))$cramer))
catcorrm(categorical,newdata)
```

### 5.1 boxplot visualiztion - pvlitM
```{r}
#anova
# pvlitM vs gender
boxplot(pvlitM~GENDER_R,data=newdata, main="gender vs pvlitM", xlab="gender", ylab="literacy score")

# pvlitM vs full/part time
boxplot(pvlitM~Full_part,data=newdata, main="pvlitM vs full/part time", xlab="full/part time", ylab="literacy score")

# pvlitM vs education
boxplot(pvlitM~ED_Level,data=newdata, main="pvlitM vs education", xlab="education", ylab="literacy score")

# pvlitM vs private/public
boxplot(pvlitM~pub_priv,data=newdata, main="pvlitM vs private/public sector", xlab="private/public sector", ylab="literacy score")

# pvlitM vs None formal education
boxplot(pvlitM~NFE12,data=newdata, main="pvlitM vs None formal education", xlab="None formal education", ylab="literacy score")

# pvlitM vs adult eduction
boxplot(pvlitM~FNFAET12,data=newdata, main="pvlitM vs adult eduction", xlab="adult eduction", ylab="literacy score")
```

### 5.2 boxplot visualiztion - pvnumM
```{r}
# pvnum vs gender
boxplot(pvnumM~GENDER_R,data=newdata, main="gender vs pvnum", xlab="gender", ylab="numeracy score")
# pvnum vs full/part time
boxplot(pvnumM~Full_part,data=newdata, main="pvnumM vs full/part time", xlab="full/part time", ylab="numeracy score")
# pvnum vs education
boxplot(pvnumM~ED_Level,data=newdata, main="pvnumM vs education", xlab="education", ylab="numeracy score")
# pvnum vs work year
boxplot(pvnumM~Years_wk_cat,data=newdata, main="pvnumM vs work year", xlab="work year", ylab="numeracy score")
```

## 3. Methods
### 3-1.Stepwise AIC & BIC(including priv&public)
```{r}
#numeracy: use data_num
full.num<-lm(pvnumM~.,data=data_num)
null.num<-lm(pvnumM~1,data=data_num)
(step(null.num,scope = list(lower=null.num,upper=full.num), direction="both", criterion 
     = "AIC", trace = F))$coefficients #ED_Level, act_lrn, Age, Gender, work_flexM,NFEHRS, public/private

(step(null.num,scope = list(lower=null.num,upper=full.num), direction="both", criterion 
     = "BIC", trace = F, k=log(nrow(data_num))))$coefficients  #ED_Level, act_lrn, Age,pub_priv, GENDER_R 

#literacture: use data_lit
full.lit<-lm(pvlitM~.,data=data_lit)
null.lit<-lm(pvlitM~1,data=data_lit)
(step(null.lit,scope = list(lower=null.lit,upper=full.lit),direction="both", criterion = 
      "AIC", trace = F))$coefficients #ED_Level, act_lrn, Age, Gender, work_flexM,NFEHRS,public/private

(step(null.lit,scope = list(lower=null.lit,upper=full.lit),direction="both", criterion = 
       "BIC", trace = F,k=log(nrow(data_lit))))$coefficients  #ED_Level, act_lrn
```

### 3-2. Stepwise AIC & BIC(exclude pub_priv)
I'll create a dataset called 'data_priv', which only contains the employees in private sector
```{r only priv}
data_priv_lit<- data_lit[data_lit$pub_priv ==1 & (!is.na(data_lit$pub_priv)), ][,-6]
data_priv_num<- data_num[data_num$pub_priv ==1 & (!is.na(data_num$pub_priv)), ][,-6]

##1)numeracy
null.priv.num <- lm(pvnumM~1,data = data_priv_num)
full.priv.num <- lm(pvnumM~.,data = data_priv_num)
library(MASS)
#AIC
(step(null.priv.num, scope = list(lower=null.priv.num,upper=full.priv.num), direction="both", criterion = "AIC",trace = F))$coefficients  
#ED_Level, act_lrn, Age, Gender, work_flexM,NFEHRS

#BIC
(step(null.priv.num, scope = list(lower=null.priv.num,upper=full.priv.num), direction="both", criterion = "BIC",k=log(nrow(data_priv_num)),trace = F))$coefficients
#ED_level, act_lrn,Age,Gender
 

##2) Literacy
null.priv.lit<- lm(pvlitM~1,data=data_priv_lit)
full.priv.lit<- lm(pvlitM~.,data=data_priv_lit)

#AIC
(step(null.priv.lit, scope = list(lower=null.priv.lit,upper=full.priv.lit), direction="both", criterion = "AIC",trace = F))$coefficients  
#ED_Level, act_lrn, Age, Gender, work_flexM,NFEHRS


#BIC
(step(null.priv.lit, scope =list(lower=null.priv.lit,upper=full.priv.lit),direction="both", criterion = "BIC",k=log(nrow(data_priv_lit)),trace = F))$coefficients
#ED_Level, act_lrn, 
```
### 3-3. Model comparison
1) check private/public sector 
```{r numeracy}
##1)numeracy:
lm.num.without<-lm(pvnumM ~ ED_Level + act_lrn + AGE_R + GENDER_R + 
    work_flexM + NFEHRS, data = data_num)
lm.num.with<-lm(pvnumM ~ ED_Level + act_lrn + AGE_R + GENDER_R + 
    work_flexM + NFEHRS + pub_priv, data = data_num)
anova(lm.num.without,lm.num.with)$`Pr(>F)`[2] 
# p-value = 0.0037**
```

```{r literacy}
##2)literacy:
lm.lit.without<-lm(formula = pvlitM ~ ED_Level + act_lrn + AGE_R+ work_flexM + 
    pub_priv + GENDER_R + NFEHRS, data = data_lit)
summary(lm.lit.without) 
lm.lit.with<-lm(formula = pvlitM ~ ED_Level + act_lrn + AGE_R + work_flexM + GENDER_R + NFEHRS, data = data_lit)

anova(lm.lit.without,lm.lit.with)$`Pr(>F)`[2] 
# p-value = 0.0185**
```

=======
2) Verify the AIC result using anova
2)-1: numeracy
```{r numeracy}
varlist <- names(data_num)[-10]
models <- lapply(varlist, function(x) {
    anova(lm(pvnumM~.,data=data_num),lm(substitute(pvnumM ~. -i, list(i = as.name(x))), data =data_num))$`Pr(>F)`[2]
})
do.call(rbind, Map(data.frame, name=as.list(varlist), pvalue=models))
```
pub_priv,AGE_R,ED_Level,work_flexM,act_lrn,FNFAET12NJR


Remark: check NFEHRS for numeracy:
From ANOVA, its corresponding p-value is 0.085; in least-squared model, its p-value is 0.137 so we would like to delete this variable
```{r NFEHRS numeracy}
summary(lm.num.with)
```


Remark: check FNFAET12NJR for numeracy
```{r}
lm.num.without2<-lm(formula = pvnumM ~ ED_Level + act_lrn + AGE_R+ work_flexM + 
     GENDER_R +pub_priv, data = data_num)

lm.num.with2<-lm(formula = pvnumM ~ ED_Level + act_lrn + AGE_R + work_flexM + GENDER_R +  pub_priv +FNFAET12NJR, data = data_num)

anova(lm.num.without2,lm.num.with2)$`Pr(>F)`[2] 
# p-value = 0.2962
```
Final model check:
```{r}
summary(lm.num.without2) #all p-value < 0.05
```

2)-1: literacy
```{r}
names(data_lit)
varlist <- names(data_lit)[-10]
models <- lapply(varlist, function(x) {
    anova(lm(pvlitM~.,data=data_lit),lm(substitute(pvlitM ~. -i, list(i = as.name(x))), data =data_lit))$`Pr(>F)`[2]
})
do.call(rbind, Map(data.frame, name=as.list(varlist), pvalue=models))
```
Variables that P-value <0.05: pub_priv,AGE_R,ED_Level,work_flexM,act_lrn

Remark: Check NFEHRS and Gender:  
```{r NEHRS literacy}
#ED_Level, act_lrn, Age, Gender, work_flexM,NFEHRS(From Stepwise AIC)
summary(lm.lit.with)
```
In this lm model, the p-value of 'GENDER_R' and 'NFEHRS' are larger than 0.05. ANOVA comparison would be used for further model diagonisis.

```{r}
#NFEHRS
anova(lm(pvlitM~ED_Level + act_lrn + AGE_R + work_flexM + NFEHRS,data=data_lit),lm(pvlitM~ED_Level + act_lrn + AGE_R + work_flexM, data=data_lit))$`Pr(>F)`[2]   #0.1554
#Gender
anova(lm(pvlitM~ED_Level + act_lrn + AGE_R + work_flexM + GENDER_R,data=data_lit),lm(pvlitM~ED_Level + act_lrn + AGE_R + work_flexM, data=data_lit))$`Pr(>F)`[2]   #0.0375
```
Final model check:
```{r}
lm_lit_final <- lm(pvlitM~ED_Level + act_lrn + AGE_R + work_flexM + GENDER_R, data=data_lit)
summary(lm_lit_final)
```

##forward AdjR^2
1) numeracy
```{r}
library(leaps)
options(digits=5)
adjr_num<-regsubsets(pvnumM~.,data=data_num,nbest=1,method = c("forward"))
summ.adjr_num<-summary(adjr_num)
which.max(summ.adjr_num$adjr2)
summ.adjr_num
# GENDER_R, AGE_R, ED_Level, pub_priv, work_flexM, act_lrn, FNFAET12NJR, NFEHRS
```


2)literacy 

```{r}
adjr_lit<-regsubsets(pvlitM~.,data=data_lit,nbest=1,method = c("forward"))
summ.adjr_lit<-summary(adjr_lit)
which.max(summ.adjr_lit$adjr2)
summ.adjr_lit
# GENDER_R, AGE_R, ED_Level, pub_priv, work_flexM, act_lrn, NFEHRS
```


