---
title: "Senior Worker 450 Code"
author: "Zhen Liu, Yutong Liu, Jingyi Huang"
date: "2018/2/1"
output: html_document
---
## Data Input
```{r data input}
s<-getwd()
substr(s, 1, nchar(s)-4)
path<-paste(substr(s, 1, nchar(s)-4),"Data/Data-NFEHRS_revised.sav",sep = "")
library(haven)
Rawdata<- read_sav(path)  #1247
```
**I think it is a nice way to start.**

## Data Pre-processing
### 1. Check Responses (Dependent variables)
```{r skil_num}
summary(Rawdata$num_use)
nrow(Rawdata[Rawdata$num_use==7,])  
```


The max of num_use is 5 so I guessed 7 may be an artefact of the data. Also, there are other four missing value in this observation.Therefore, we would exclude this obervation(num_use=7).
```{r qqplot }
newdata<- Rawdata[Rawdata$num_use!=7,]  #1246
par(mfrow=c(2,2))
#1)skill_num
summary(newdata$num_use)
qqnorm(newdata$num_use)
qqline(newdata$num_use)
#2) skill_lit
summary(newdata$liter_use)
qqnorm(newdata$liter_use)
qqline(newdata$liter_use)
#3) test_num
summary(newdata$pvnumM)
qqnorm(newdata$pvnumM)
qqline(newdata$pvnumM)
#4) test_lit
summary(newdata$pvlitM)
qqnorm(newdata$pvlitM)
qqline(newdata$pvlitM)
```
** It would be nice to make a table or data frame containing all the summary information just using one summary fucntion, such as summary(newdata)[,c(15,16,17,18)] . It would be nicer to add the title for each plot :)**
#summary(newdata$num_use)
The top left and top right Q-Q plot are not on a straight line. We might need some further transformation on the responses(num_use, liter_use)
 
The bottom left and bottom right Q-Q plot seem on a straight line so it is fair to say that the two responses(pvnumM, pvlitM) of our sample are normal distributed. The assumption of linear regression is statisfied.

### 2.Show summary statistics and check Missing Value
```{r missing value}
sum.data<-summary(newdata)[7,]
sum.data[!is.na(sum.data)]
```
** Wow This is very nice! interesting codes for me...**
There are 501 missing value in the 'Mgr' column and 1015 missing value in the 'Mgr_c' colum, which contain both more than 40 % of the total number of the observations. I would suggest that progressing our regression analysis without these two variables first. There are 39 missing values in the 'pub_priv' and we don't know which sector they are.

### 3.Check Independent Variable
```{r independent} 
##1.ED_Level
unique(newdata$ED_Level)

table(newdata$ED_Level) #only one observation =8 and only 4 observations=3; change them to 3
newdata<- within(newdata, ED_Level[ED_Level==8]<-3 )
newdata<- within(newdata, ED_Level[ED_Level==4]<-3 )

##2.Categorical variable
newdata$pub_priv <- as.factor(newdata$pub_priv)
newdata$GENDER_R <- as.factor(newdata$GENDER_R)
newdata$ED_Level <- as.factor(newdata$ED_Level)
newdata$Full_part <- as.factor(newdata$Full_part)
newdata$NFE12 <- as.factor(newdata$NFE12)
newdata$FNFE12JR <- as.factor(newdata$FNFE12JR)
newdata$FNFAET12NJR <- as.factor(newdata$FNFAET12NJR)
newdata$FNFAET12JR <- as.factor(newdata$FNFAET12JR)

##3. "FNFAET12" "FNFAET12JR" "FNFAET12NJR" t-test:
#1)pvlitM
t.test(pvlitM~FNFAET12, data = newdata)$p.value #p-value < 2.2e-16
t.test(pvlitM~FNFAET12JR, data = newdata)$p.value #p-value = 9.813e-13
t.test(pvlitM~FNFAET12NJR, data = newdata)$p.value #p-value = 0.0001315
#2)pvnumM
t.test(pvnumM~FNFAET12, data = newdata)$p.value #p-value < 2.2e-16
t.test(pvnumM~FNFAET12JR, data = newdata)$p.value #p-value < 2.2e-16
t.test(pvnumM~FNFAET12NJR, data = newdata)$p.value #p-value = 6.66e-06
##they are all statistically significant and FNFAET12 is the union of other two variables.

##4.FNFE12JR and FNFAET12JR
table(newdata$FNFE12JR == newdata$FNFAET12JR) #They are exactly the same: use one of them
```
**-There is some typos on the first comment. When I checked it, only 4 observations have value 4.
-Maybe it could be a good idea to show the union relationship between the three variables like this: 
table(newdata\$FNFAET12==(newdata\$FNFAET12JR+newdata$FNFAET12NJR))
- for the ##2 where you change the data type as factor, I think we can use simpler codes such as :newdata["pub_priv","GENDER_R","ED_Level","Full_part","NFE12","FNFE12JR","FNFAET12NJR","FNFAET12JR"] <- lapply(newdata["pub_priv","GENDER_R","ED_Level","Full_part","NFE12","FNFE12JR","FNFAET12NJR","FNFAET12JR"] , factor)
**
### 4. Two dataset:
```{r two dataset}
#1) proficiency test score(numeracy) + possbile factor
data_num<- newdata[,c(1,2,5,7,9,10,13,14,17,19,20,23,24,25)]
data_num<- na.omit(data_num) #1207
#2) proficiency test score(literacy) + possbile factor
data_lit<- newdata[,c(1,2,5,7,9,10,13,14,17,18,20,23,24,25)]
data_lit<- na.omit(data_lit) #1207
```

### 5. Correlations
```{r correlations}
newdata.numeric<-newdata[,c("AGE_R","Years_wk","work_flexM","work_lrnM","act_lrn","NFEHRS","pvlitM","pvnumM")]
m<-cor(newdata.numeric)
if (!require('corrplot')) install.packages('corrplot')
library('ggplot2')
library(vcd)
corrplot(m, method = "circle") #plot matrix
categorical<-c("GENDER_R","Full_part","ED_Level","pub_priv","NFE12","FNFAET12JR","FNFAET12NJR")
catcorrm <- function(categorical, newdata) sapply(categorical, function(y) sapply(categorical, function(x) 
assocstats(table(unlist(newdata[,x]), unlist(newdata[,y])))$cramer))
catcorrm(categorical,newdata) #only FNFAET12JR and NFE12 have very strong correlation(0.83376057)
```
** We may need this too: if (!require('vcd')) install.packages('vcd') **
### 5.1 boxplot visualiztion - pvlitM
```{r}
#anova
# pvlitM vs gender
boxplot(pvlitM~GENDER_R,data=newdata, main="gender vs pvlitM", xlab="gender", ylab="literacy score")

# pvlitM vs full/part time
boxplot(pvlitM~Full_part,data=newdata, main="pvlitM vs full/part time", xlab="full/part time", ylab="literacy score")

# pvlitM vs education
boxplot(pvlitM~ED_Level,data=newdata, main="pvlitM vs education", xlab="education", ylab="literacy score")

# pvlitM vs private/public
boxplot(pvlitM~pub_priv,data=newdata, main="pvlitM vs private/public sector", xlab="private/public sector", ylab="literacy score")

# pvlitM vs None formal education
boxplot(pvlitM~NFE12,data=newdata, main="pvlitM vs None formal education", xlab="None formal education", ylab="literacy score")

# pvlitM vs adult eduction
boxplot(pvlitM~FNFAET12,data=newdata, main="pvlitM vs adult eduction", xlab="adult eduction", ylab="literacy score")
```
**I think this code is fine. But if I have to give you a suggestion, I would recommend you to use for loop not to repeat the same codes.**
### 5.2 boxplot visualiztion - pvnumM
```{r}
# pvnum vs gender
boxplot(pvnumM~GENDER_R,data=newdata, main="gender vs pvnum", xlab="gender", ylab="numeracy score")
# pvnum vs full/part time
boxplot(pvnumM~Full_part,data=newdata, main="pvnumM vs full/part time", xlab="full/part time", ylab="numeracy score")
# pvnum vs education
boxplot(pvnumM~ED_Level,data=newdata, main="pvnumM vs education", xlab="education", ylab="numeracy score")
# pvnum vs work year
```
**I think this is fine. We only have 3 box plots.**
## Methods
### 1. Literacy
#### 1-1 Forward Adjusted R^2
```{r}
library(leaps)
options(digits=5)
adjr_lit<-regsubsets(pvlitM~.,data=data_lit,nbest=1,nvmax = 14)
?regsubsets
summ.adjr_lit<-summary(adjr_lit)
summ.adjr_lit$adjr2
summ.adjr_lit$cp
which.max(summ.adjr_lit$adjr2)#8
which.min(summ.adjr_lit$cp)#8

par(mfrow=c(1,2))
plot(adjr_lit, scale="adjr2")
plot(adjr_lit, scale="Cp")



summ.adjr_lit
# GENDER_R, AGE_R, ED_Level(3), pub_priv, work_flexM, act_lrn, NFEHRS
```
**When I run the 'which.max(summ.adjr_lit$adjr2)', I got 8. I am not sure about what 7 means...**
#### 1-2 Stepwise AIC 
```{r}
table(newdata$pub_priv)
library(MASS)
full.lit<-lm(pvlitM~.,data=data_lit)
null.lit<-lm(pvlitM~1,data=data_lit)
(step(null.lit,scope = list(lower=null.lit,upper=full.lit),direction="forward", trace = F))$coefficients #ED_Level, act_lrn, Age, Gender, work_flexM,NFEHRS,public/private
#GENDER_R + AGE_R + ED_Level + pub_priv + work_flexM + act_lrn + NFEHRS, same as above
```

#### 1-3 Stepwise AIC (only private)
```{r }
data_priv_lit<- data_lit[data_lit$pub_priv ==1 & (!is.na(data_lit$pub_priv)), ][,-6]
null.priv.lit<- lm(pvlitM~1,data=data_priv_lit)
full.priv.lit<- lm(pvlitM~.,data=data_priv_lit)
(step(null.priv.lit, scope = list(lower=null.priv.lit,upper=full.priv.lit), direction="forward", criterion = "AIC",trace = F))$coefficients   #ED_Level, act_lrn, Age, Gender, work_flexM, NFEHRS
```

#### 1-4 Model Comparison
1) check private/public sector 
```{r lit_priv/pub}
lm.lit.without<-lm(formula = pvlitM ~ ED_Level + act_lrn + AGE_R+ work_flexM + 
    pub_priv + GENDER_R + NFEHRS, data = data_lit)
lm.lit.with<-lm(formula = pvlitM ~ ED_Level + act_lrn + AGE_R + work_flexM + GENDER_R + NFEHRS, data = data_lit)
anova(lm.lit.without,lm.lit.with)$`Pr(>F)`[2]  # p-value = 0.020435**
summary(lm.lit.with) #p-value of NFEHRS = 0.147
```

2) ANOVA: With vs without one variable
```{r anova}
varlist <- names(data_lit)[-10]
models <- lapply(varlist, function(x) {
    anova(lm(pvlitM~.,data=data_lit),lm(substitute(pvlitM ~. -i, list(i = as.name(x))), data =data_lit))$`Pr(>F)`[2]
})
do.call(rbind, Map(data.frame, name=as.list(varlist), pvalue=models))
```
Variables that P-value <0.05: pub_priv,AGE_R,ED_Level,work_flexM,act_lrn
**This code is fine. You may also use drop1() function, just for your information.**

3)
```{r compare}
anova(lm(pvlitM~ED_Level + act_lrn + AGE_R + work_flexM + NFEHRS,data=data_lit),lm(pvlitM~ED_Level + act_lrn + AGE_R + work_flexM, data=data_lit))$`Pr(>F)`[2]   #NFEHRS: 0.12197
anova(lm(pvlitM~ED_Level + act_lrn + AGE_R + work_flexM + GENDER_R,data=data_lit),lm(pvlitM~ED_Level + act_lrn + AGE_R + work_flexM, data=data_lit))$`Pr(>F)`[2]   #Gender: 0.04734
```

### 2. Numeracy
#### 2-1 Forward Adjusted R^2
```{r}
adjr_num<-regsubsets(pvnumM~.,data=data_num,nbest=1,method = c("exhaustive"),nvmax=16)
summ.adjr_num<-summary(adjr_num)
which.max(summ.adjr_num$adjr2)
plot(adjr_num,scale = "adjr2")
# GENDER_R, AGE_R, ED_Level, pub_priv, work_flexM, act_lrn,NFE121,FNFAET12JR1， FNFAET12NJR1,NFEHRS
```

#### 2-2 Stepwise AIC
```{r}
full.num<-lm(pvnumM~.,data=data_num)
null.num<-lm(pvnumM~1,data=data_num)
(step(null.num,scope = list(lower=null.num,upper=full.num), direction="forward", criterion 
     = "AIC", trace = F))$coefficients #ED_Level, act_lrn, Age, Gender, public/private, work_flexM, NFEHRS
step(full.num,scope = list(lower=null.num), direction="backward", criterion 
     = "AIC", trace = F)
```

#### 2-3. Stepwise AIC (Only private)
```{r only priv}
data_priv_num<- data_num[data_num$pub_priv ==1 & (!is.na(data_num$pub_priv)), ][,-6]
null.priv.num <- lm(pvnumM~1,data = data_priv_num)
full.priv.num <- lm(pvnumM~.,data = data_priv_num)
(step(null.priv.num, scope = list(lower=null.priv.num,upper=full.priv.num), direction="forward", criterion = "AIC",trace = F))$coefficients  
#ED_Level, act_lrn, Age, Gender, work_flexM,NFEHRS
```

#### 2-4. Model comparison
1) check private/public sector 
```{r priv/pub}
lm.num.without<-lm(pvnumM ~ ED_Level + act_lrn + AGE_R + GENDER_R + 
    work_flexM + NFEHRS, data = data_num)
lm.num.with<-lm(pvnumM ~ ED_Level + act_lrn + AGE_R + GENDER_R + 
    work_flexM + NFEHRS + pub_priv, data = data_num)

summary(lm.num.with) #NFEHRS:0.13733
anova(lm.num.without,lm.num.with)$`Pr(>F)`[2] # p-value = 0.0037**
```

2) ANOVA: With vs without one variable
```{r anova}
varlist <- names(data_num)[-10]
models <- lapply(varlist, function(x) {
    anova(lm(pvnumM~.,data=data_num),lm(substitute(pvnumM ~. -i, list(i = as.name(x))), data =data_num))$`Pr(>F)`[2]
})
do.call(rbind, Map(data.frame, name=as.list(varlist), pvalue=models))
```
pub_priv,AGE_R,ED_Level,work_flexM,act_lrn,FNFAET12NJR


3)
```{r compare}
anova(lm(formula = pvnumM ~ ED_Level + act_lrn + AGE_R+ work_flexM + GENDER_R +pub_priv, data = data_num),lm(formula = pvnumM ~ ED_Level + act_lrn + AGE_R + work_flexM + GENDER_R +  pub_priv +NFEHRS, data = data_num))$`Pr(>F)`[2] #NFEHRS:0.13303

anova(lm(formula = pvnumM ~ ED_Level + act_lrn + AGE_R+ work_flexM + GENDER_R +pub_priv, data = data_num),lm(formula = pvnumM ~ ED_Level + act_lrn + AGE_R + work_flexM + GENDER_R +  pub_priv +FNFAET12NJR, data = data_num))$`Pr(>F)`[2]  #FNFAET12NJR: 0.31301

anova(lm(formula = pvnumM ~ ED_Level + act_lrn + AGE_R+ work_flexM + GENDER_R +pub_priv, data = data_num),lm(formula = pvnumM ~ ED_Level + act_lrn + AGE_R + work_flexM + GENDER_R +  pub_priv +NFE12, data = data_num))$`Pr(>F)`[2] #NFE12: 0.45681

anova(lm(formula = pvnumM ~ ED_Level + act_lrn + AGE_R+ work_flexM + GENDER_R +pub_priv, data = data_num),lm(formula = pvnumM ~ ED_Level + act_lrn + AGE_R + work_flexM + GENDER_R +  pub_priv +FNFAET12JR, data = data_num))$`Pr(>F)`[2]  #FNFAET12JR: 0.7498

anova(lm(formula = pvnumM ~ ED_Level + act_lrn + AGE_R+ work_flexM + GENDER_R +pub_priv, data = data_num),lm(formula = pvnumM ~ ED_Level + act_lrn + AGE_R + work_flexM + GENDER_R +  pub_priv + Years_wk, data = data_num))$`Pr(>F)`[2] #Years_work: 0.40395

anova(lm(formula = pvnumM ~ ED_Level + act_lrn + AGE_R+ work_flexM + GENDER_R +pub_priv, data = data_num),lm(formula = pvnumM ~ ED_Level + act_lrn + AGE_R + work_flexM + GENDER_R +  pub_priv +FNFAET12JR + NFE12 + FNFAET12NJR , data = data_num))$`Pr(>F)`[2]  # 0.16506

```

4)Final model check:
```{r final_lit}
summary(lm(formula = pvnumM ~ ED_Level + act_lrn + AGE_R+ work_flexM + 
     GENDER_R +pub_priv, data = data_num)) #all p-value < 0.05
```


